---
title: "Analyzing news from the Guardian"
author: "Sofia Gil-Clavel"
date: "2026-04-13"
output: 
  md_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,eval = FALSE)
```

This section is based on:

Gil-Clavel, S. (2025). R-Course: R for Social Scientists (Version 1.0) [Data set]. https://github.com/SofiaG1l/R_Course/. doi: https://doi.org/10.5281/zenodo.17184812.


# Data Analysis

Now that we have annotated all the data, let's analyse it. 

First, let's clean the environment. So that, we are not carrying any variables 

```{r clean environment}

rm(list=ls())
gc()

```

For this section, we will use the following packages:

```{r load packages}

library(tidyverse)
library(tidytext)
library(SnowballC)
library(hunspell)

```


Now, let's open your data.

```{r open the data}

load("~/All_Annotated_Data.RData")


```


## Homogenizing the data

The first step in any text analysis is to homogenize the data. This means performing:

1. Tokenizing
2. Stemming
3. Stopword removal

First, to tokenize the data means to turn the sentences into some unit of analysis. For example, characters, words, n-grams (i.e., two or more words), sentences, etc. To tokenize the words, we will use the function `unnest_tokens()` from the package `tidytext`:

```{r tokenize}

TXT=All_Annotated_Data%>%
  unnest_tokens(words, text)

```

If you check the data, now there is a new column called **word**. We will now continue working with that column.

Second, stopword removal refers to deleting all those words that are too common to provide any context, e.g., the, a, those, etc...

We can remove them using the `stop_words` data frame from the package `tidytext`:

```{r stopword removal}

TXT=TXT%>%
  filter(!stemms%in%stop_words$word)

```


Third, stemming, as some words are just conjugations of a verb, then we can try to consider them as the same word. This is called stemming.

There are multiple algorithms to perform stemming. Today, we will focus on two of them `snowball` and `hunspell`.

To use snowball, you just need to call the function `wordStem()` from the package `SnowballC`:

```{r stemming using snowball}
TXT=TXT%>%
  mutate(stem_snow = SnowballC::wordStem(words))

```

To use hunspell, you just need to call the function `hunspell_stem()` from the package `hunspell`:

```{r stemming using hunspell}

TXT=TXT%>%
  mutate(stem_huns = hunspell::hunspell_stem(words))

# We need to unnest the vector values
TXT=TXT%>%
  unnest_wider(stem_huns, simplify = FALSE,names_sep = "_")

TXT=TXT%>%
  mutate(stem_huns = ifelse(!is.na(stem_huns_2),stem_huns_2,stem_huns_1))%>%
  # remove the other columns
  select(-stem_huns_1,-stem_huns_2)

```

Now, you can open the data to check the new columns. Which algorithm performed the best? 

Craete a new column called **stemm** with the result you found most useful, if any, and delete the unused columns.

```{r stemming}

TXT=TXT%>%
  mutate(stemms=stem_huns)%>%
  select(-stem_snow,-stem_huns,-stem_huns_3)

# Some of the words in "stemms" are NAs. So, let's replace those with the original word:

TXT<-TXT%>%
  mutate(stemms=ifelse(is.na(stemms),words,stemms))%>%
  select(-words)

```


We can also consider removing numbers and other special characters, as they may not contribute any special information to our study:

```{r only characters}

TXT<-TXT%>%
  filter(str_detect(stemms,"[:alpha:]"))

```


## Text Mining

### Frequencies

Now, we can look for some interesting patterns. Let's do this by counting the number of words per document and year.

```{r frequencies}

TXT<-TXT%>%
  group_by(url,datetime,headline,stemms)%>%
  summarise(total=n())

```

Let's visualize the words that appear the most:

```{r Viz Wordcloud}

TXT%>%
  ungroup()%>%
  group_by(stemms)%>%
  summarise(total=sum(total))%>%
  slice_max(order_by = total,n=30)%>%
  mutate(angle = 45 * sample(-2:2, n(), replace = TRUE, 
                             prob = c(1, 1, 4, 1, 1)))%>%
  ggplot(aes(label = stemms, size = total, angle = angle)) +
  ggwordcloud::geom_text_wordcloud_area(
    rm_outside = TRUE,
    area_corr = TRUE) + # 
  scale_size_area(max_size = 20) +
  theme_minimal()

```
Let's visualize the frequencies by article:

```{r Viz Freq 1}

TXT%>%
  slice_max(order_by = total,n=10,with_ties = FALSE)%>%
  ggplot(aes(total,reorder_within(stemms,total,headline)))+
  geom_col()+
  scale_y_reordered()+
  facet_wrap(str_trunc(headline,25)~.,scales="free")+
  theme_light()

```


### TF-IDF

Sometimes frequency alone is not enough to have a good understanding of what each document is about. A better option is to use the `tf-idf`. This value considers how relevant the term is in comparison with other terms in other documents. term frequency and inverse document frequency.

```{r tf-idf}

TXT=TXT%>%
  bind_tf_idf(stemms, url, total)

```

Let's visualize the tf-idf values:

```{r Viz2}

TXT%>%
  slice_max(order_by = tf_idf,n=10,with_ties = FALSE)%>%
  ggplot(aes(tf_idf,reorder_within(stemms,tf_idf,headline)))+
  geom_col()+
  scale_y_reordered()+
  facet_wrap(str_trunc(headline,25)~.,scales="free")+
  theme_light()

```

