---
title: "Using httr2"
author: "Sofia Gil-Clavel"
date: "2026-05-18"
output: 
  md_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,eval = FALSE)

```

# Data retrieval

In this section of the tutorial, you will learn how to download data from the 
Guardian. The assumption is that you already requested and obtained your developer The Guardian key. You should have done it through this link: [https://open-platform.theguardian.com/access/](https://open-platform.theguardian.com/access/).

This part of the tutorial uses the packages:

* tidyverse
* httr2

So, before starting please make sure that you have already installed them.

Now, let's load the packages into R:

```{r load packages}

library(tidyverse)
library(httr2)

```


## Accessing data via an API

In general, an Application Programming Interface (API) is used for computer programs to communicate with each other. This part of the tutorial is based on [Use APIs to source your data in R](https://raw.githack.com/Delft-RCafe/resources/main/themes/apis_in_r/apis-in-r.html) \(2023\) by Bjørn Bartholdy & Aleksandra Wilczynska.

The most commonly use API type is the Representational State Transfer (REST). The REST API uses the HTTP protocol to send and receive standardized responses to a server. 

There are five HTTP methods that you can use when making an API request to a server:

* **GET**: Retrieves data.
* **POST**: Creates a new record.
* **PUT**: Modifies or replaces an entire record.
* **PATCH**: Modifies or updates parts of a record.
* **DELETE**: Deletes an entire record.

Besides those methods, the other important components are:

* **HTTP method**: Explains what action you will perform.
* **endpoint**: The URL to find the resource you are trying to reach on the internet.
* **headers**: Provides information relevant both for us and for the server. Such as your The Guardian key!
* **body**: These are specific parameters that the server requires.

### Using the package `httr2`

The package `httr2` can be used together with the `tidyverse` package. 

The **endpoint** to access The Guardian news is [https://content.guardianapis.com/search](https://content.guardianapis.com/search). We are going to save this URL in a variable called `req`:

```{r endpoint}

req<-request("https://content.guardianapis.com/search")

```

Now, we need to:

1. Set the HTTP method to **GET**. This is done with the function `req_method()`.
2. Pass our The Guardian credentials via the **header** using the function `req_headers()`. Please go ahead and replace the variable `KEY` with your own key. 
3. Pass other parameters through the **body** via `req_url_query()`. The other parameters are:

  * **Query:** For the query, you can check the possible format here [https://open-platform.theguardian.com/documentation/](https://open-platform.theguardian.com/documentation/).
  * **Other parameters**: You can find a list of the parameters here [https://open-platform.theguardian.com/documentation/search](https://open-platform.theguardian.com/documentation/search). 


```{r body}

query="(artificial intelligence) OR (chat-gpt) OR (open AND ai) OR (llm) OR (large language model*)"

resp <- req |>
  req_method("GET") |> # request method
  req_headers("Accept" = "application/json",
              "api-key" = KEY) %>%
  req_url_query("q" = query,
                "from-date" = "2014-01-01",
                "to-date" = "2014-06-01")%>%
  req_perform()

```

If you explore the variable `resp`, you will see that it is a very long string. To be able to read it, we will first turn it into JSON format:

```{r response}

json_response <- resp_body_string(resp)

```


Now, we can turn the `json` format into a data frame using the function `fromJSON()` from the same package. Let's also turn this data frame into a tibble:

```{r from json to data frame}

resp_df <- jsonlite::fromJSON(json_response, flatten = TRUE)

TheGuardianAI<-tibble(resp_df$response$results)

```


## Saving your results

Now, we can save your text as a tibble:

```{r save results, eval=FALSE}

save(TheGuardianAI,file = "TheGuardianAI_2025.RData")

```


# Creating a bigger corpus

As time is precious, we will create a bigger corpus by each of us retrieving values for different years.

I will tell each of you which year to retrieve and where to store the data.


